A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-09 text/manuscript.Rmd
Update filtering of genes method

One of the limitations of knowledge-based pathway analysis is that it heavily relies on 
experimental results for the curation and validation of its pathways. The resulting 
pathways are therefore not specific to a particular tissue type but provides information
for all tissue type. Before utilizing the gene to pathway lookup table, we considered the
impact of deleterious passenger mutations for a given cancer (non-specific genes that are
not expressed) as they may introduce biases in our pathway analysis. We addressed this by
identifying genes that are lowly/not expressed in a tissue type. Using the TCGA RNA 
sequencing data adjusted for batch effect dataset (https://pancanatlas.xenahubs.net), we 
filtered the genes and tumor samples that were not included in our analysis, grouped the 
tumor samples by tissue type and computed the mean expression value for each gene. A minimum
threshold of 10 transcripts per million was set for expressed genes based on an inflection
point observed when plotting the mean expression values of genes ranked by increasing 
expression values for each tissue type. Genes that did not meet this threshold were 
considered not expressed. This operation produced a lookup table for gene expression
status in each tumor sample for 18,127 genes.

>>>>

We filtered likely erroneous mutations due to transcription coupled repair. Our approach
was to determine the status of each gene (expressed or not expressed) per tissue type in
order to exclude gene mutations that we have evidence for no expression. To do this, we 
obtained the TCGA RNA sequencing data adjusted for batch effect dataset (https://pancanatlas.xenahubs.net).
Using the data, we removed the genes and tumor samples that were not included in our analysis, 
grouped the tumor samples by tissue type and computed the mean expression value for each 
gene. A minimum threshold of 10 transcripts per million was set for expressed genes based 
on an inflection point observed when plotting the mean expression values of genes ranked 
by increasing expression values for each tissue type. Genes that did not meet this threshold 
were considered not expressed. This operation produced a lookup table for gene expression
status in each tumor sample for 18,127 genes.
